---
title: "Forest Height Map with City Structures"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_notebook
---

# Forest Height and Urban Structure Map for Ruhrgebiet

This notebook combines forest height data with OpenStreetMap urban infrastructure 
to create a comprehensive visualization of the region.

## Setup

```{r setup, message=FALSE, warning=FALSE}
# Load utility functions
source("utils.R")

# Load required libraries
load_libs(c(
  "tidyverse",
  "sf",
  "geodata",
  "terra",
  "classInt",
  "osmdata",
  "purrr",
  "rayshader"
))

# Set stable Overpass server
set_overpass_url("https://overpass.kumi.systems/api/interpreter")
```

## Configuration

```{r config}
area <- "bochum"  # Options: "bochum" or "ruhrgebiet"
data_dir <- "data"
if (!dir.exists(data_dir)) dir.create(data_dir)
```

## 1. Download and Prepare Administrative Boundaries

```{r boundaries}
# Get country borders
country_borders <- get_country_borders("DEU", 3)

# Filter for North Rhine-Westphalia
nrw <- country_borders %>%
  filter(NAME_1 == "Nordrhein-Westfalen")

# Define Ruhrgebiet region
ruhrgebiet <- nrw %>%
  filter(
    NAME_2 %in% c(
      "Bochum", "Bottrop", "Dortmund", "Duisburg", "Essen", "Gelsenkirchen",
      "Hagen", "Hamm", "Herne", "Mülheim an der Ruhr", "Oberhausen",
      "Recklinghausen", "Unna", "Wesel", "Ennepe-Ruhr-Kreis"
    )
  )

# Filter for Bochum
bochum <- nrw %>% filter(NAME_2 == "Bochum")

# Select area based on configuration
if (area == "bochum") {
  sel <- bochum
} else if (area == "ruhrgebiet") {
  sel <- ruhrgebiet
} else {
  stop("area must be 'bochum' or 'ruhrgebiet'")
}

# Create bounding box
bbox <- st_bbox(sel)
bbox <- c(bbox["xmin"], bbox["ymin"], bbox["xmax"], bbox["ymax"])

cat("Selected area:", area, "\n")
cat("Bounding box:", bbox, "\n")
```

## 2. Download Forest Height Data

```{r forest-download}
# Forest height data URL
urls <- c("https://libdrive.ethz.ch/index.php/s/cO8or7iOe5dT2Rt/download?path=%2F3deg_cogs&files=ETH_GlobalCanopyHeight_10m_2020_N51E006_Map.tif")

dest_files <- c("ETH_GlobalCanopyHeight_10m_2020_N51E006_Map.tif")
full_dest_files <- file.path(data_dir, dest_files)
force_download <- FALSE

# Download files if needed
for (i in seq_along(urls)) {
  dest_path <- full_dest_files[i]
  if (!file.exists(dest_path) || isTRUE(force_download)) {
    message("Downloading: ", dest_files[i])
    download.file(urls[i], destfile = dest_path, mode = "wb", quiet = FALSE)
  } else {
    message("Already present, skipping download: ", dest_path)
  }
}

raster_files <- full_dest_files[file.exists(full_dest_files)]
if (length(raster_files) == 0) {
  stop("No raster files found in ", data_dir)
}
```

## 3. Process Forest Height Data

```{r forest-processing}
# Union geometry for cropping
region_union <- sel %>% st_union()

# Load and crop raster files
forest_height_list <- lapply(raster_files, terra::rast)

forest_height_rasters <- lapply(
  forest_height_list,
  function(x) {
    terra::crop(
      x,
      terra::vect(region_union),
      snap = "in",
      mask = TRUE
    )
  }
)

# Mosaic if multiple rasters
if (length(forest_height_rasters) > 1) {
  forest_height_mosaic <- do.call(terra::mosaic, forest_height_rasters)
} else {
  forest_height_mosaic <- forest_height_rasters[[1]]
}

# Aggregate to reduce resolution (faster processing)
forest_height_region <- forest_height_mosaic %>%
  terra::aggregate(fact = 10)

# Convert to dataframe for ggplot
forest_height_df <- forest_height_region %>%
  as.data.frame(xy = TRUE)
names(forest_height_df)[3] <- "height"

cat("Forest height data processed\n")
cat("Dimensions:", nrow(forest_height_df), "pixels\n")
```

## 4. Prepare Bounding Box Tiles (for large areas)

```{r bbox-tiles}
split_bbox <- function(bbox) {
  xm <- (bbox["xmin"] + bbox["xmax"]) / 2
  ym <- (bbox["ymin"] + bbox["ymax"]) / 2
  
  list(
    c(bbox["xmin"], bbox["ymin"], xm, ym),
    c(xm, bbox["ymin"], bbox["xmax"], ym),
    c(bbox["xmin"], ym, xm, bbox["ymax"]),
    c(xm, ym, bbox["xmax"], bbox["ymax"])
  )
}

if (area == "bochum") {
  tiles <- list(bbox)  # No split needed
} else {
  tiles <- split_bbox(bbox)  # Split for larger area
}
```

## 5. Download OSM Infrastructure Data

```{r osm-download, message=FALSE, warning=FALSE}
# Highways (major roads)
highways <- get_osm_cached(
  paste0("data/highways_", area, ".rds"),
  function() {
    combine_osm(
      lapply(tiles, function(t)
        safe_osm(
          t,
          "highway",
          c(
            "motorway", "trunk", "primary", "secondary",
            "tertiary", "motorway_link", "trunk_link"
          )
        )
      )
    )
  }
)

# Streets (residential and minor roads)
streets <- get_osm_cached(
  paste0("data/streets_", area, ".rds"),
  function() {
    combine_osm(
      lapply(tiles, function(t)
        safe_osm(
          t,
          "highway",
          c("residential", "living_street", "service", 
            "unclassified", "pedestrian")
        )
      )
    )
  }
)

# Waterways
water <- get_osm_cached(
  paste0("data/water_", area, ".rds"),
  function() {
    combine_osm(
      lapply(tiles, function(t)
        safe_osm(t, "waterway", c("river", "stream", "canal"))
      )
    )
  }
)

# Water bodies
water_bodies <- get_osm_cached(
  paste0("data/waterbodies_", area, ".rds"),
  function() {
    combine_osm(
      lapply(tiles, function(t)
        safe_osm(t, "water", c("river", "lake", "reservoir", "pond"))
      )
    )
  }
)

# Water areas
water_areas <- get_osm_cached(
  paste0("data/waterareas_", area, ".rds"),
  function() {
    combine_osm(
      lapply(tiles, function(t)
        safe_osm(t, "natural", "water")
      )
    )
  }
)

cat("OSM data downloaded successfully\n")
```

## 6. Define Color Schemes

```{r colors}
# Forest height colors
forest_cols <- c(
  "white", "#ffd3af", "#fbe06e",
  "#6daa55", "#205544"
)

forest_texture <- colorRampPalette(
  forest_cols,
  bias = 2
)(6)

# Urban infrastructure colors
color_roads <- rgb(0.42, 0.449, 0.488)
color_water <- rgb(0.2, 0.4, 0.6, 0.5)

# Calculate breaks for forest height
breaks <- classInt::classIntervals(
  forest_height_df$height,
  n = 5,
  style = "quantile"
)$brks
```

## 7. Create Combined Map

```{r combined-map, fig.width=12, fig.height=14}
combined_map <- ggplot() +
  # Forest height base layer
  geom_tile(
    data = forest_height_df,
    aes(x = x, y = y, fill = height)
  ) +
  scale_fill_gradientn(
    name = "Height (m)",
    colors = forest_texture,
    breaks = round(breaks, 0)
  ) +
  # Water features
  geom_sf(
    data = water_areas$osm_polygons,
    fill = color_water,
    col = color_water,
    alpha = 0.6,
    lwd = 0,
    inherit.aes = FALSE
  ) +
  geom_sf(
    data = water_bodies$osm_polygons,
    fill = color_water,
    col = color_water,
    alpha = 0.6,
    lwd = 0,
    inherit.aes = FALSE
  ) +
  geom_sf(
    data = water$osm_lines,
    col = color_water,
    linewidth = 0.8,
    alpha = 0.7,
    inherit.aes = FALSE
  ) +
  # Road network
  geom_sf(
    data = streets$osm_lines,
    col = color_roads,
    linewidth = 0.3,
    alpha = 0.5,
    inherit.aes = FALSE
  ) +
  geom_sf(
    data = highways$osm_lines,
    col = color_roads,
    linewidth = 0.6,
    alpha = 0.8,
    inherit.aes = FALSE
  ) +
  # Administrative boundary
  geom_sf(
    data = sel,
    fill = NA,
    color = "black",
    linewidth = 0.5,
    inherit.aes = FALSE
  ) +
  # Coordinate system and limits
  coord_sf(
    xlim = c(bbox[1], bbox[3]),
    ylim = c(bbox[2], bbox[4]),
    expand = FALSE,
    crs = 4326
  ) +
  guides(
    fill = guide_legend(
        direction = "vertical",
        keyheight = unit(5, "mm"),
        keywidth = unit(5, "mm"),
        title.position = "top",
        label.position = "right",
        title.hjust = .5,
        label.hjust = .5,
        ncol = 1,
        byrow = F
    )
) +
  theme_minimal() +
  theme(
      axis.line = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      legend.position = "right",
      legend.title = element_text(
          size = 11, color = "grey10"
      ),
      legend.text = element_text(
          size = 10, color = "grey10"
      ),
      panel.grid.major = element_line(
          color = "white"
      ),
       panel.grid.minor = element_line(
          color = "white"
      ),
      plot.background = element_rect(
          fill = "white", color = NA
      ),
      legend.background = element_rect(
          fill = "white", color = NA
      ),
      panel.border = element_rect(
          fill = NA, color = "white"
      ),
      plot.margin = unit(
          c(
              t = 0, r = 0,
              b = 0, l = 0
          ), "lines"
      )
  )
  # Legend
  # guides(
  #   fill = guide_colorbar(
  #     direction = "vertical",
  #     barheight = unit(50, "mm"),
  #     barwidth = unit(5, "mm"),
  #     title.position = "top",
  #     title.hjust = 0.5,
  #     label.hjust = 0.5
  #   )
  # ) +
  # # Theme
  # theme_void() +
  # theme(
  #   legend.position = "right",
  #   legend.title = element_text(size = 11, color = "grey10"),
  #   legend.text = element_text(size = 10, color = "grey10"),
  #   plot.background = element_rect(fill = "white", color = NA),
  #   legend.background = element_rect(fill = "white", color = NA),
  #   plot.margin = unit(c(0, 0, 0, 0), "lines")
  # ) +
  # labs(
  #   title = paste("Forest Height and Urban Infrastructure -", 
  #                 tools::toTitleCase(area))
  # )

print(combined_map)
```

## 8. Create a version without a legend

```{r no-legend-map}
combined_map_no_legend <- combined_map + 
  theme(legend.position = "none")

#print(combined_map_no_legend)
```

## 9. Create a 3D Version of the map
```{r calculate width and height}
h <- nrow(forest_height_region)
w <- ncol(forest_height_region)

rayshader::plot_gg(
    ggobj = combined_map,
    width = w / 1000,
    height = h / 1000,
    scale = 150,
    solid = F,
    soliddepth = 0,
    shadow = T,
    shadow_intensity = .99,
    offset_edges = F,
    sunangle = 315,
   # window.size = c(800, 800),
    zoom = .4,
    phi = 30,
    theta = -30,
    multicore = T
)

rayshader::render_camera(
    phi = 50,
    zoom = .7,
    theta = 45
)



```

```{r export, eval=FALSE}
# Export settings
dpi <- 300
width_inch <- 16.5
height_inch <- 23.4

# Save the map
ggsave(
  filename = paste0("output/forest_city_map_", area, ".png"),
  plot = combined_map,
  width = width_inch,
  height = height_inch,
  dpi = dpi,
  bg = "white"
)

cat("Map exported successfully\n")
```

## Summary

This notebook creates a comprehensive map combining:

- **Forest height data** from ETH Global Canopy Height dataset (2020)
- **Road network** from OpenStreetMap (highways and streets)
- **Water features** (rivers, streams, lakes, water bodies)
- **Administrative boundaries** for the selected region

The visualization shows the spatial relationship between natural forest cover 
and urban development patterns in the region.